Learning \strips\ action models with classical planning

### 2017-11-01 doc/icaps-18/strips-learning.tex
updated formalization pages 1 and 2

\title{Learning STRIPS action models with classical planning}

@@

\maketitle
\begin{abstract}
This paper presents a novel approach for learning STRIPS action models from examples that compiles this particular inductive learning task into a classical planning task. Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given. What is more, the compilation accepts partially specified action models and can also be used to validate whether certain observations of plan executions follow a given STRIPS action model (even if the action model is not fully specified).
\end{abstract}

@@

On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. The application of inductive ML to the learning of STRIPS action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
\begin{itemize}
\item The inputs to ML algorithms (the learning data) usually are finite vectors that encode the value of fixed features for a given set of objects. The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
\item The traditional output of ML algorithms is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks). When learning STRIPS action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of the corresponding planning domain.
\end{itemize}

The learning of STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans. Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation, so an off-the-shelf planner can address the learning task. The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
\begin{enumerate}
\item Is flexible to different amounts of available input knowledge. The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models. In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given STRIPS action model.
\end{enumerate}

The paper is organized as follows. The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the STRIPS action model (the output of the addressed learning task). The second section formalizes the learning of STRIPS action models with regard to different amounts of available input knowledge. The third and fourth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to constrain the space of possible action models and make learning more practicable. Finally, the last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.

\section{Background}
This section defines the planning models used on this work as well as the STRIPS action model, the output of the learning task addressed in this paper.

\subsection{Classical planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$. We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

@@

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of {\em effects} $\eff(a)\in\mathcal{L}(F)$. An action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.

@@

\subsection{Classical planning with conditional effects}
Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects. Conditional effects allow us to compactly define actions whose particular effects depend on the current state. Nowadays many classical planners cope with conditional effects without compiling them away. In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.

Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of conditional effects $\cond(a)$. Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the condition, and $E\in\mathcal{L}(F)$, the effect.

An action $a\in A$ is {\em applicable} in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
\[
\eff(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
\]
i.e.~effects whose conditions hold in $s$. The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.

\subsection{The STRIPS action model and the {\em variable} objects}
This work addresses the learning of action schemes defined in the STRIPS language~\cite{fikes1971strips}. To formalized the output of this learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that the fluents in $F$ are instantiated from these predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.

Formally, each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

Likewise, we assume that STRIPS actions are instantiated from STRIPS operator schemes. Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
\begin{figure}
\begin{footnotesize}
\begin{verbatim}
(:action stack
  :parameters (?x1 ?x2 - object)
  :precondition (and (holding ?x1)
                     (clear ?x2))
  :effect (and (not (holding ?x1))
               (not (clear ?x2))
               (handempty)
               (clear ?x1)
               (on ?x1 ?x2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of a STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.}
\label{fig:stack}
\end{figure}

Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and whose cardinality, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$. For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because operators {\small\tt stack} and {\small\tt unstack}, with the maximum arity, have two parameters each.

Let us define also a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$. In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.

We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and a enumeration of the variable objects bound by the operator arity, $pars(\xi)\in\Omega_v^{ar(\xi)}$. The headers for a 4-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

\section{Learning STRIPS action models}
Learning STRIPS action models from fully available input knowledge, i.e. from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward. In this case, a STRIPS operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions. Likewise, the preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.

@@

 \caption{\small Example of a task for learning a STRIPS action model in the blocksworld.}
\label{fig:lexample}
\end{figure}

A solution to a learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.

@@

In this case, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples). We assume however, that the headers of the operators schema are known.

\section{Learning STRIPS action models with planning}
Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects. The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).  For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$. We call this process the validation of the programmed STRIPS action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}. If information about the corresponding plan $\pi_t\in \Pi$ is available, is used in the validation.
\end{enumerate}

@@

Now we are ready to formalize our compilations for learning STRIPS action models with classical planning. We start with $\Lambda'$ because this learning task requires smaller input knowledge. Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
\begin{itemize}
\item $F_{\Lambda'}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$). If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
\end{itemize}
\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
\item $G_{\Lambda'}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
\item $A_{\Lambda'}$ contains actions of three different kinds:
\begin{enumerate}
\item The actions for programming an operator schema $\xi\in\Xi$, which includes:
\begin{itemize}
\item The actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.

@@

The compilation is {\em complete} in the sense that it does not discard any possible STRIPS action model.

@@

\section{Constraining the learning hypothesis space with additional input knowledge}
Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of STRIPS action models more practicable.

@@

\subsection{Previously specified action models}
Our compilation does not require to start learning STRIPS action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.

@@

In the extreme, when a fully specified STRIPS action model is given, the compilation validates whether an observed plan follows the given model. In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given STRIPS action model is {\em valid} for the given set of examples. If $P_{\Lambda'}$ is unsolvable it means that the given STRIPS action model is invalid since it is not compliant with all the given examples. Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.

@@

Back in the 90's various systems were aimed at learning operators mostly via interaction with the environment. The proposal of the LIVE system was to capture and formulate observable features of objects and use them to acquire and refine operators \cite{ShenS89}. OBSERVER, on the other hand, updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}. These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge. The recent work reported in \cite{WalshL08} shows that although learning STRIPS operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.

Whilst the aforementioned works deal with full state observability, action model learning has also been studied in domains
where there is partial or missing state observability~\cite{jimenez2012review}. The {\sf ARMS} approach~\cite{yang2007learning}, which works when no or partial intermediate states are given, extracts frequent action sets from plans that share a common set of parameters and frequent literal-action pairs with the help of the initial and goal state of the plans. The extracted information is used then to define a set of weighted constraints that must hold for all the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver. Since complex planning problems may result in a too large MAX-SAT representation to be solved efficiently, {\sf ARMS} implements a hill-climbing method that models the actions approximately. Consequently, the output model of {\sf ARMS} may be inconsistent with the examples. The approach presented in \cite{amir:alearning:JAIR08} deals with partial observability. Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation
of observations by models of actions through a CNF formula. The learning algorithm updates the formula of the belief state with every action and observation in the sequence. This update makes sure that the new formula represents all the transition relations consistent with the actions and observations. The formula returned at the end includes all consistent models, which can then be retrieved with additional processing. Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations. An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then STRIPS rules are derived from the classifiers' parameters.

@@

\section{Conclusions}
The paper presented a novel approach for learning STRIPS action models from examples exclusively using classical planning. The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.

@@

The size of the compiled classical planning instances depends on the number of input examples. The empirical results show that our approach can generate non-trivial models from very small data sets. In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon. The same happens with all the actions for programming an action effect can also be applied in a single planning step.

>>>>

\title{Learning \strips\ action models with classical planning}

@@

\maketitle
\begin{abstract}
This paper presents a novel approach for learning \strips action models from examples that compiles this particular inductive learning task into classical planning. Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given). What is more, the compilation accepts partially specified action models and can be used to validate whether certain observations of plan executions follow a given \strips\ action model, even if this model is not fully specified.
\end{abstract}

@@

On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. The application of inductive ML to the learning of \strips\ action models~\cite{fikes1971strips}, the vanilla action model for planning, is not straightforward though:
\begin{itemize}
\item The inputs to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects. The input for learning planning action models are observations of plan executions (where each plan possibly has different length).
\item The traditional output of ML algorithms is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks). When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
\end{itemize}

The learning of \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans. Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation. The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
\begin{enumerate}
\item Is flexible to different amounts of available input knowledge. The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models. In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given \strips\ action model.
\end{enumerate}

The paper is organized as follows. The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the \strips\ action model (the output of the addressed learning task). The second section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge. The third and fourth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to constrain the space of possible action models and make learning more practicable. Finally, last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.

\section{Background}
This section defines the planning models used on this work and the output of the learning task addressed in the paper.

\subsection{Classical planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

@@

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ comprises three sets of literals:
\begin{itemize}
\item $\pre(a)$, called {\em preconditions}, that defines the literals that must hold for the action to be applicable.
\item $\eff^+(a)$, called {\em positive effects}, that defines the fluents set to true by the action application.
\item $\eff^-(a)$, called {\em negative effects}, that defines the fluents set to false by the application of the action. 
\end{itemize}
We say that an action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

@@

\subsection{Classical planning with conditional effects}
Our approach for leaning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects. Conditional effects allow us to compactly define actions whose effects depend on the current state. Many classical planners cope with conditional effects without compiling them away. In fact, supporting conditional effects is a requirement of IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.

Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of {\em conditional effects} $\cond(a)$. Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.

An action $a\in A$ is {\em applicable} in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
\[
triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
\]
i.e.~effects whose conditions hold in $s$.

The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\del(s,a))\cup\add(s,a)\}$ where $del(s,a)\subseteq triggered(s,a)$ and $add(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects .

\subsection{The \strips\ action schema and the {\em variable} objects}
This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}. Figure~\ref{fig:stack} shows the schema that corresponds to the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.

\begin{figure}
\begin{scriptsize}
\begin{verbatim}
(:action stack
  :parameters (?x1 ?x2 - object)
  :precondition (and (holding ?x1) (clear ?x2))
  :effect (and (not (holding ?x1)) (not (clear ?x2))
               (handempty) (clear ?x1) (on ?x1 ?x2)))
\end{verbatim}
\end{scriptsize}
 \caption{\small Example of a \strips\ operator schema that corresponds to the {\em stack} action from the {\em blocksworld} represented in PDDL.}
\label{fig:stack}
\end{figure}

To formalize the output of the learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that fluents $F$ are instantiated from these predicates, as in PDDL. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$. Likewise, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes.

Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and that is bound by the maximum arity of an action in a given planning frame. For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.

Let us define also a new set of fluents $F_{v}$, s.t. $F\cap F_v=\emptyset$, that results instantiating $\Psi$ using only {\em variable objects} in $\Omega_v$. This set defines the elements that can appear in an action schema. In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.

We are now ready to define $\Xi$, a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$, an enumeration of the variable objects bound by the operator arity. The headers for a 4-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

\section{Learning \strips action models}
Learning \strips action models from fully available input knowledge, i.e. from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward. In this case, a \strips operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions. Likewise, the preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.

@@

 \caption{\small Example of a task for learning a \strips action model in the blocksworld.}
\label{fig:lexample}
\end{figure}

A solution to a learning task $\Lambda$ is a set of \strips operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.

@@

In this case, a solution must not only determine a possible \strips action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples). We assume however, that the headers of the operators schema are known.

\section{Learning \strips action models with planning}
Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects. The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the \strips action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed \strips action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).  For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$. We call this process the validation of the programmed \strips action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}. If information about the corresponding plan $\pi_t\in \Pi$ is available, is used in the validation.
\end{enumerate}

@@

Now we are ready to formalize our compilations for learning \strips action models with classical planning. We start with $\Lambda'$ because this learning task requires smaller input knowledge. Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
\begin{itemize}
\item $F_{\Lambda'}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$). If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips operator schema $\xi\in \Xi$.
\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
\end{itemize}
\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
\item $G_{\Lambda'}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
\item $A_{\Lambda'}$ contains actions of three different kinds:
\begin{enumerate}
\item The actions for programming an operator schema $\xi\in\Xi$, which includes:
\begin{itemize}
\item The actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.

@@

The compilation is {\em complete} in the sense that it does not discard any possible \strips action model.

@@

\section{Constraining the learning hypothesis space with additional input knowledge}
Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips action models more practicable.

@@

\subsection{Previously specified action models}
Our compilation does not require to start learning \strips action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.

@@

In the extreme, when a fully specified \strips action model is given, the compilation validates whether an observed plan follows the given model. In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips action model is {\em valid} for the given set of examples. If $P_{\Lambda'}$ is unsolvable it means that the given \strips action model is invalid since it is not compliant with all the given examples. Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.

@@

Back in the 90's various systems were aimed at learning operators mostly via interaction with the environment. The proposal of the LIVE system was to capture and formulate observable features of objects and use them to acquire and refine operators \cite{ShenS89}. OBSERVER, on the other hand, updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}. These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge. The recent work reported in \cite{WalshL08} shows that although learning \strips operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.

Whilst the aforementioned works deal with full state observability, action model learning has also been studied in domains
where there is partial or missing state observability~\cite{jimenez2012review}. The {\sf ARMS} approach~\cite{yang2007learning}, which works when no or partial intermediate states are given, extracts frequent action sets from plans that share a common set of parameters and frequent literal-action pairs with the help of the initial and goal state of the plans. The extracted information is used then to define a set of weighted constraints that must hold for all the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver. Since complex planning problems may result in a too large MAX-SAT representation to be solved efficiently, {\sf ARMS} implements a hill-climbing method that models the actions approximately. Consequently, the output model of {\sf ARMS} may be inconsistent with the examples. The approach presented in \cite{amir:alearning:JAIR08} deals with partial observability. Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation
of observations by models of actions through a CNF formula. The learning algorithm updates the formula of the belief state with every action and observation in the sequence. This update makes sure that the new formula represents all the transition relations consistent with the actions and observations. The formula returned at the end includes all consistent models, which can then be retrieved with additional processing. Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations. An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.

@@

\section{Conclusions}
The paper presented a novel approach for learning \strips action models from examples exclusively using classical planning. The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.

@@

The size of the compiled classical planning instances depends on the number of input examples. The empirical results show that our approach can generate non-trivial models from very small data sets. In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the \strips actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon. The same happens with all the actions for programming an action effect can also be applied in a single planning step.
